---
title: 'EDS 221 Day 6: Activity'
author: "Takeen Shamloo"
date: "2024-08-19"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Day 6: Activity

## Task 2 ----

```{r}
# Load Packages ----
library(tidyverse)
library(here)
library(janitor)
```

## Read/Process Data

```{r}
lobsters <- read_csv(here("data","Lobster_Abundance_All_Years_20210412.csv"), na = c("-99999", "")) |>
  clean_names() |>
  uncount(count) # Uses the number in the 'count' column to determine how many times it will duplicate that row of data.
  
head(lobsters)
```

## Find counts and mean sizes by site & year

```{r}
lobster_summary <- lobsters |>
  dplyr::group_by(site, year) |>
  summarize(total_count = n(),
            mean_carapace_len = mean(size_mm, na.rm = TRUE))
```

### Plot summarized data by site and year

```{r}
ggplot(lobster_summary, aes(x = year, y = total_count, colour = site)) +
  geom_point() +
  geom_line() + 
  facet_wrap(~ site) +
  labs(x = "Year", y = "Lobster Count", title = "Lobster Count per Year")
```

## Find the proportion of legal lobsters at each site in 2020

```{r}
legal_lobsters_2020 <- lobsters |>
  filter(year == 2020) |>
  mutate(legal = case_when(
    size_mm < 79.76 ~ FALSE,
    size_mm >= 79.76 ~ TRUE)) |>
  group_by(site, legal) |>
  summarize(total_count = n()) 
```

### Plot proportion of legal lobsters using stacked column graph

```{r}
ggplot(legal_lobsters_2020, aes(x = site, y = total_count, fill = legal)) + 
  geom_col(position = "fill") + 
  labs(y = "Proportion", x = "Site", fill = "Legal Status") +
  theme_minimal()
```

NAPL and IVEE had the largest proportion of legal lobsters in 2020.

## Task 3 ----

### filter() practice ----
```{r}
ex_a <- lobsters |>
  filter(site == c("IVEE", "CARP", "NAPL"))

ex_b <- lobsters |>
  filter(month == 8)

ex_c <- lobsters |>
  filter(site == "AQUE" | size_mm > 70)

ex_d <- lobsters |>
  filter(site != "NAPL")
```

### group_by() practice ----

```{r}
ex_e <- lobsters |>
  group_by(site) |>
  summarize(mean_len = mean(size_mm, na.rm = TRUE),
            std_len = sd(size_mm, na.rm = TRUE))

ex_f <- lobsters |>
  group_by(site, month) |>
  summarise(max_len = max(size_mm, na.rm = TRUE))
```

### mutate() practice

```{r}
ex_g <- lobsters |>
  mutate(size_cm = size_mm/10)

ex_h <- lobsters |>
  mutate(site = tolower(site))

ex_i <- lobsters |>
  mutate(area = as.character(area))
class(ex_i$area) 
typeof(ex_i$area)
```

### case_when() practice ----

```{r}
ex_j <- lobsters |>
  mutate(size_bin = case_when(
    size_mm <= 70 ~ "small",
    size_mm > 70 ~ "large"
  ))
head(ex_j$size_bin)

ex_k <- lobsters |>
  mutate(designation = case_when(
    (site == "IVEE" | site == "NAPL") ~ "MPA",
    !(site == "IVEE" | site == "NAPL") ~ "not MPA"
  ))
head(ex_k$designation)
tail(ex_k$designation)
```

```




